### 2020-03-19 files_from_roberts_thesis/results/clustering/convae.tex

\section{Deep clustering algorithms}\label{sec:convae_clustering}

In the previous section we demonstrated a potential for clustering of events from an AT-TPC experiment. To build on this result we will in this section explore the application of the mixture of autoencoders (MIXAE) algorithm introduced in section \ref{sec:deep_clustering}. The other algorithm introduced in section \ref{sec:deep_clustering}, deep clustering with convolutional autoencoders (DCEC), consistently collapsed to just one cluster for all datasets.

In the MIXAE algorithm the hyper-parameters to adjust are all the ordinary parameters that we introduced in table \ref{tab:convae_hyperparams}. In addition to those parameters we have the weighting of the loss terms: $\theta$, $\alpha$ and $\gamma$. These weighting parameters are attached to the reconstruction loss, sample entropy and batch-wise entropy respectively. \cite{Zhang} note that these parameters are very important for the model performance and so we focus primarily on these.

We empirically freeze the convolutional autoencoder hyperparameters to compress the original $128 \times 128$ to a $8 \times 8$ pixel grid using four convolutional layers. The parameters chosen for the autoencoders are listed in full in table \ref{tab:mixe_ae_hyperparams}.

\begin{table}[H]
\renewcommand*{\arraystretch}{0.5}
\centering
\caption{Hyperparameters selected for the autoencoder components of the MIXAE algorithm, see table \ref{tab:convae_hyperparams} for a full description of the parameters.}\label{tab:mixe_ae_hyperparams}
\setlength{\extrarowheight}{15pt}
\hspace*{-0.5in}
\begin{tabular}{ll}
\toprule
Hyperparameter & Value \\
\midrule
\multicolumn{2}{l}{Convolutional parameters: } \\
\midrule
Number of layers & $4$ \\
Kernels & $[3,\,3,\,3,\,3]$\\
Strides & $[2,\,2,\,2,\,2]$ \\
Filters & $[64,\, 32, \,16, \,8,]$ \\ 
\midrule
\multicolumn{2}{l}{Network parameters: } \\
\midrule
Activation & LReLu \\
Latent type & None \\
Latent dimension & 20  \\
$\beta$ & N/A \\
Batchnorm & False \\
\midrule
\multicolumn{2}{l}{Optimizer parameters: } \\
\midrule
$\eta$ & $10^{-3}$ \\
$\beta_1$ & $0.9$ \\
$\beta_2$ & $0.99$ \\
\bottomrule
\end{tabular}
\end{table}

Since there are then only three remaining hyperparameters we choose to perform a coarse grid-search as described in section \ref{sec:hyperparam_search_arch}.

\begin{table}
\centering
\caption{Hyperparameter grid for the MIXAE loss weighting terms. The grid is given as exponents for logarithmic scales.}\label{tab:mixae_loss_weights}
\begin{tabular}{lll}
\toprule
Parameter & Grid & Scale \\
\midrule 
$\theta$ & $[-1,\, 5]$ & Logarithmic \\
$\alpha$ & $[-5,\, -1]$ & Logarithmic \\
$\gamma$ & $[3,\, 5]$ & Logarithmic
\end{tabular}
\end{table}

\subsection{Simulated AT-TPC data}

To train the MIXAE clustering algorithm, we use the large simulated dataset with $M=80000$ points, evenly distributed between proton- and carbon-events. The algorithm is trained on a subset of $60000$ of these samples, and we track performance on the remaining $20000$ events.

The grids selected for the search are listed in table \ref{tab:mixae_loss_weights}. The search yielded an optimal configuration with

\begin{align}
\theta = 10^{-1}, \\
\alpha = 10^{-2}, \\
\gamma = 10^5.
\end{align}

Finally, for these parameters we re-ran the algorithm $N=10$ times to investigate the stability of the algorithm. The results are reported in figure \ref{fig:mixae_sim}. We observe that while the algorithm can achieve very strong performance, with an adjusted rand index (ARI) $ > 0.8$, it fluctuates strongly with repeated experiments. The $ARI$ is a clustering performance measure similar to accuracy, but adjusted for chance. We elaborate on the $ARI$ score in section \ref{sec:unsupervised_perf}. As mentioned in section \ref{sec:mixae} the batch-entropy has a second minimum when the cluster confidences are near equal. It is this behavior we observe in figure \ref{fig:mixae_sim}.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/sim_mixae.pdf}
\caption[Clustering performance of MIXAE on simulated AT-TPC data]{Performance for the MIXAE model on simulated AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run. }\label{fig:mixae_sim}
\end{figure}

\subsection{Filtered AT-TPC data}

We repeat the optimization steps in the previous section for the filtered AT-TPC data. The exception being that we allow the algorithm to train on the labelled samples. Beginning with a wide grid equal to the grid used for the simulated data we searched over all parameter configurations to find promising values. We then performed a local search around these values to pin-point the hyperparameter configuration. This search yielded the optimal hyper-parameters

\begin{align}
\theta &= 10^{1}, \\
\alpha &= 10^{-1}, \\
\gamma &= 3.162\times 10^3.
\end{align}

\noindent The results of the runs are included in figure \ref{fig:mixae_clean}. We observe ghat the highest performing models reach an $ARI > 0.5$, which is higher than the performance achieved by the K-means algorithm applied to the VGG16 latent space.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/clean_mixae.pdf}
\caption[Clustering performance of MIXAE on filtered event AT-TPC data]{Performance for the MIXAE model on filtered AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_clean}
\end{figure}

\subsection{Full AT-TPC data}

As in the two previous sections, we repeat the same procedure of iterative grid searches on the MIXAE loss-weights. Each configuration is re-run a total of $N=10$ times to capture fluctuations in the performance before a final selection is made on the hyperparameters. For the full dataset the MIXAE hyperparameters converge to the same values as for the clean data, i.e.

\begin{align}
\theta &= 10^{1}, \\
\alpha &= 10^{-1}, \\
\gamma &= 3.162\times 10^3.
\end{align}

\noindent As with the previous datasets we include a plot of the loss curves and performance measures for $N=10$ runs with the same loss-weight parameters, shown in figure \ref{fig:mixae_full}.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{./plots/real_mixae.pdf}
\caption[Clustering performance of MIXAE on full event AT-TPC data]{Performance for the MIXAE model on un-filtered AT-TPC data. In the top row the loss components are plotted for each run, and in the bottom row the adjusted rand index (ARI) and clustering accuracy are shown. Each run is color-coded with the ARI achieved at the end of the run.}\label{fig:mixae_full}
\end{figure}

\subsection{Comparing performance}

It is also interesting to compare and contrast the clustering results from the MIXAE model with those of the VGG16$+$K-means outside the fairly abstract accuracies and rand scores. It is especially interesting to compare the cluster assignments, as they can inform further research efforts in the clustering of AT-TPC events. We illustrate the clustering with confusion matrices that are shown in figure \ref{fig:mixae_confmat}. From these matrices, we observe that the MIXAE applied to the clean data correctly clusters the noise events. Additionally, it identifies two proton clusters. We observe that these proton clusters are both less pure than the VGG16+K-means clusters and that there does not seem to be a visually meaningful difference between these clusters. The latter is inferred from figure \ref{fig:filtered_mixae_clster_repr}.

Applied to the real data the MIXAE correctly separates the proton class, however it is unable to separate the carbon events from the amorphous noise events or from the proton cluster.

\begin{figure}[H]
\centering

	\subfloat{
	\includegraphics[width=0.5\textwidth]{plots/filtered_mixae_conf_mat.pdf}
}
	\hspace{-1cm}
	\subfloat{
	\includegraphics[width=0.5\textwidth]{plots/full_mixae_conf_mat.pdf}
}
\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on filtered and full AT-TPC events. The true labels indicate samples belonging to the p (proton), carbon (C), or other classes. }\label{fig:mixae_confmat}
\end{figure}

Lastly we wish to further investigate if there are systematic differences between proton events that were placed in different clusters for the clean and full data. From figure \ref{fig:mixae_confmat} we see that the MIXAE algorithm creates two proton clusters for the filtered data, and places about fifty per-cent of the proton events in a cluster with the amorphous "other" events. We extract some proton events from these clusters to inspect whether systematic differences occur.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{plots/clean_cluster_repr.pdf}
\caption[Selection of carbon events in differing clusters]{Selection of carbon events belonging to different clusters. Each row represents one of the clusters for the filtered data as shown in figure \ref{fig:mixae_confmat}. There seems to be no clear distinction between these rows.}\label{fig:filtered_mixae_clster_repr}	
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{plots/real_cluster_repr.pdf}	
\caption[Selection of proton events in differing clusters]{Selection of proton events belonging to different clusters. Each row represents one of the clusters for the full data as shown in figure \ref{fig:mixae_confmat}. We observe a tendency for the more noisy events to be placed in the "other" cluster}\label{fig:full_mixae_clster_repr}	
\end{figure}