### 2020-06-07 doc/Manuscript/plots/mixae_clustering.tex
added table to mixae

\begin{tabular}{lllll}
\toprule
{} & \multicolumn{2}{c}{Accuracy} &   \multicolumn{2}{c}{ARI} \\
\midrule
{} & Top 1 & $\mu \pm \sigma$ & Top 1 & $\mu \pm \sigma$ \\
Simulated &  0.96 &  $0.74 \pm 0.16 $  & 0.84 &  $0.33 \pm 0.32 $\\
Filtered  & 0.75 &  $0.71 \pm 0.04 $ & 0.52 &  $0.38 \pm 0.14 $\\
Full      & 0.71 &  $0.61 \pm 0.07 $ &  0.32 &  $0.09 \pm 0.10 $   \\
\bottomrule
\end{tabular}