### 2020-08-04 doc/Manuscript/plots/mixae_clustering.tex
Updates from Overleaf

\begin{tabular}{lllll}
\toprule
{} & \multicolumn{2}{c}{Accuracy} &   \multicolumn{2}{c}{ARI} \\
\midrule
{} & {\color{red}Top 1}{\color{blue} init $= 10$} &  {\color{red}$\mu \pm \sigma$}{\color{blue} init $= 1$} & Top 1 & $\mu \pm \sigma$ \\
Simulated &  0.96{\color{blue} $\pm 0.00$} &  $0.74 \pm 0.16 $  & 0.84 &  $0.33 \pm 0.32 $\\
Filtered  & 0.75{\color{blue} $\pm 0.00$} &  $0.71 \pm 0.04 $ & 0.52 &  $0.38 \pm 0.14 $\\
Raw & 0.71{\color{blue} $\pm 0.00$} &  $0.61 \pm 0.07 $ &  0.32 &  $0.09 \pm 0.10 $   \\
\bottomrule
\end{tabular}