### 2021-03-13 credit_statement.tex
v2 - revision submission

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}
{\bf Robert Solli:} Conceptualization, Methodology, Software, Validation, Formal Analysis, Writing - Original Draft, Writing - Review \& Editing, Visualization,
{\bf Daniel Bazin:} Conceptualization, Investigation, Writing - Original Draft, Writing - Review \& Editing, Supervision,
{\bf Morten Hjorth-Jensen:} Resources, Writing - Original Draft, Writing - Review \& Editing, Supervision
{\bf Michelle Kuchera:} Conceptualization, Methodology, Resources, Writing - Original Draft, Writing - Review \& Editing, Supervision,
{\bf Ryan Strauss:} Methodology, Software, Validation, Formal Analysis, Writing - Original Draft, Writing - Review \& Editing,

\end{document}